\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual All-Loop Integrands

### 2023-09-05 notes/2loopNLSM.tex
Added some color and a few equation labels.

Even beginning at one loop, we encounter classes of cuts that are
difficult to properly define due to hidden non-physical singularities.
The quintessential example of such a ``pathological'' cut is the 3-1
two-particle cut at one loop which hides a bubble-on-external-leg
singularity
\begin{equation}
  \begin{tikzpicture}
\begin{feynman}
\vertex (a1) at (-.8, -0.6) {1};
\vertex (a2) at (-.8, 0.6) {3};
\vertex (a3) at (-1, 0) {2};
\vertex [dot, scale=2](mid1) at (0,0){};
\vertex [dot, scale=1.5,hgrey0](mid2) at (0,0){};
\vertex [dot, scale=2](mid3) at (1,0){};
\vertex [dot, scale=1.5,hgrey0](mid4) at (1,0){};
\vertex (a4) at (1.8, 0) {4};
\diagram{
(mid1) --[ultra thick,](a1),
(mid1) --[ultra thick,](a2),
(mid1) --[ultra thick,](a3),
%(mid1) --[ultra thick,](mid3),
(mid1) --[ultra thick,out=60,in=120,min distance=0.4cm](mid3),
(mid1) --[ultra thick,out=-60,in=-120,min distance=0.4cm](mid3),
(mid3) --[ultra thick](a4),
};
\end{feynman}
\end{tikzpicture}
\ni
\frac{\delta(\textcolor{cut-color}{k_4^2})}{\textcolor{ucp-color}{s\, k_4^2}}
n \left[
  \begin{tikzpicture}
    \ver{e1}{-1,0};
    \ver{e2}{-0.6,0.4};
    \ver{e3}{-0.3,0.4};
    \ver{i1}{-0.6,0};
    \ver{i2}{-0.3,0};
    \ver{i3}{0,0};
    \ver{i4}{0.8,0};
    \ver{co}{$(i3)!0.5!(i4)$};
    \ver{e4}{1.2,0};
    \draw[ultra thick] (e1)--(i1)--(i2)--(i3) to [out=80,in=100,min distance=0.4cm](i4)--(e4);
    \draw[ultra thick,ucp-color](i1)--(i2)--(i3);
    \draw[ultra thick,cut-color](i4)--(e4);
    \draw[ultra thick] (i3) to [out=-80,in=-100,min distance=0.4cm](i4);
    \draw[ultra thick] (e2) -- (i1);
    \draw[ultra thick] (e3)--(i2);
    \draw[ultra thick,dashed,cut-color] (co)--++(90:0.5);
    \draw[ultra thick,dashed,cut-color] (co)--++(-90:0.5);
  \end{tikzpicture}
  \right]
\end{equation}
in which one of the uncut propagators is set equal to an external
on-shell propagator.  Discussions about how to resolve these types of
singularities can be found in Refs.~\cite{Edison:2022jln,
  Edison:2022smn, Edison:2020uzf, Geyer:2015bja, Bern:2012uf}
\citepls{Minahan}.  In addition to the bubble-on-external-leg cuts,
there is a related class of cuts in which an uncut propagator is set
to zero by an \emph{internal cut condition}, for instance,
\begin{equation}
  \label{eq:other-path}
  \begin{tikzpicture}
    \pgfmathsetmacro{\ri}{1};
    \pgfmathsetmacro{\re}{0.6};
    \pgfmathsetmacro{\ang}{35};
    \begin{feynman}
      \blob{i1}{\ang:\ri};
      \blob{i2}{180-\ang:\ri};
      \blob{i3}{180+\ang:\ri};
      \blob{i4}{-\ang:\ri};
      \blob{i5}{ $(i1)!0.5!(i2)$ };
      \ver{e1}{45:\ri+\re};
      \ver{e2}{135:\ri+\re};
      \ver{e3}{225:\ri+\re};
      \ver{e4}{-45:\ri+\re};

@@

Ignoring tadpoles and BEL graphs, there are 14 cubic 4pt 2-loop graphs (see later picture) which are related to each other via 21 Jacobi relations.
As mentioned in (review section), color-kinematics duality ensures that the numerator of every graph can be expressed in terms of a basis of the double box and penta-triangle,
\be 
\NLSMBasisDoubleBox \text{ and } \NLSMBasisPentaTriangle .
\ee
For a scalar theory like NLSM the numerator only depends on dot products of momenta.
The basis of Mandelstam invariants for both of these graphs can be taken to be
\be
\{ k_{13}, k_{23}, k_{15}, k_{16}, k_{25}, k_{26}, k_{35}, k_{36}, k_{55}, k_{56}, k_{66} \},
\ee
where $k_{ij}$ means $k_i \cdot k_j$, $k_5=\ell_1$, and $k_5 = \ell_2$.
%%%%
% Basis for both diagrams
% {d[p[1], p[3]], d[p[1], p[6]], d[p[1], p[7]], d[p[2], p[3]],  d[p[2], p[6]], d[p[2], p[7]], d[p[3], p[6]], d[p[3], p[7]],  d[p[6], p[6]], d[p[6], p[7]], d[p[7], p[7]]}
%%%%
Every vertex in NLSM scales as $k^2$ so the numerator of each diagram scales as $k^{12}$.
At this mass dimension there are 8,008 independent dot products of momenta.
With two basis graphs there is a total of 16,016 ansatz parameters.\footnote{\texttt{Mathematica} is not suited to equations with sixteen thousand parameters so a different method is desirable.
One possibility is to use a solver designed for integral reduction.
However, these solvers typically assume that the matrix has a very low density which is not the case for equations coming from NLSM which typically have a density of a few percent.
While the equations in YM naturally separate into different ``helicity'' sectors (or $e_i \cdot e_j$ and $e_i \cdot k_j$ sectors in $D$ dimensions), there is no such separation of sectors in a scalar theory where any product of momenta can interfere with any other.
The NLSM constraint equations were solved using a custom solver, \texttt{FiniteFieldSolve}, designed to solve large linear systems of arbitrary density.
\texttt{FiniteFieldSolve} will be released publicly shortly.}
Imposing all of the Jacobi relations eliminates all but 4,473 of the original free parameters.
Further imposing graph symmetries reduces the number of free parameters to 1,243.

@@

We follow the same general procedure as outlined above, again
identifying the double-box and penta-triangle as the Jacobi basis
graphs and building a generic ansatz for each of their numerators.
Because we are now considering pure Yang--Mills, each monomial in the
local ansatz must consist of five Lorentz scalar dot products instead
of the six for NLSM, and each monomial must be linear in each of the
four external gluon polarizations.  Thus, the ansatz will be built
from terms of the form
\begin{align}
  \pol_1 \cdot k_2, \quad \pol_1 \cdot k_3, \quad& \pol_1 \cdot \ell_1, \quad \pol_1 \cdot \ell_2, \notag \\
  \pol_2 \cdot k_1, \quad \pol_2 \cdot k_3, \quad& \pol_2 \cdot \ell_1, \quad \pol_2 \cdot \ell_2, \notag \\
  \pol_3 \cdot k_1, \quad \pol_3 \cdot k_2, \quad& \pol_3 \cdot \ell_1, \quad \pol_3 \cdot \ell_2,  \\
  \pol_4 \cdot k_1, \quad \pol_4 \cdot k_2, \quad& \pol_4 \cdot \ell_1, \quad \pol_4 \cdot \ell_2, \notag \\
  \pol_1 \cdot \pol_2, \quad \pol_1 \cdot \pol_3, \quad& \pol_1 \cdot \pol_4, \quad \pol_2 \cdot \pol_3, \notag \\
  \pol_2 \cdot \pol_4, \quad& \pol_3 \cdot \pol_4  \notag
\end{align}
in addition to the purely momentum-dependent terms show in \sect{}.
In our labeling of the two loop Jacobi-basis diagrams, shown in
\ace{Figure out fig}, we find that the power counting bound proposed
by Bern, Davies, and Nohle leads to an initial ansatz with 10,000
terms in the double-box, and 9849 terms for the penta-triangle.
Without power counting restrictions, both the double-box and the
penta-triangle ansatz have an ansatz with 10,010 terms each.  Imposing
diagram automorphisms and maximal-cut gauge invariance on the two
basis diagrams, we reduce the number of terms to XX for the
double-box, and YY for the penta-triangle.  The power-counting
restriction of BDN would further reduce these counts to 2232 and 4072.
In Ref.~\cite{Bern:2015ooa}, the unitarity matching constraints were
imposed using a spanning set of minimal physical cuts, shown in
\cref{fig:ym-spanning}.  Instead of also directly going to the
spanning cuts, we work within the framework of the Method of Maximal
Cuts \citepls{} in order to identify the simplest cut that is in
tension with the kinematic Jacobi relations.

>>>>

Even beginning at one loop, we encounter classes of cuts that are
difficult to properly define due to hidden non-physical singularities.
The quintessential example of such a ``pathological'' cut is the 3-1
two-particle cut at one loop which hides a bubble-on-external-leg
singularity
\begin{equation}
  \begin{tikzpicture}
\begin{feynman}
\vertex (a1) at (-.8, -0.6) {1};
\vertex (a2) at (-.8, 0.6) {3};
\vertex (a3) at (-1, 0) {2};
\vertex [dot, scale=2](mid1) at (0,0){};
\vertex [dot, scale=1.5,hgrey0](mid2) at (0,0){};
\vertex [dot, scale=2](mid3) at (1,0){};
\vertex [dot, scale=1.5,hgrey0](mid4) at (1,0){};
\vertex (a4) at (1.8, 0) {4};
\diagram{
(mid1) --[ultra thick,](a1),
(mid1) --[ultra thick,](a2),
(mid1) --[ultra thick,](a3),
%(mid1) --[ultra thick,](mid3),
(mid1) --[ultra thick,out=60,in=120,min distance=0.4cm](mid3),
(mid1) --[ultra thick,out=-60,in=-120,min distance=0.4cm](mid3),
(mid3) --[ultra thick](a4),
};
\end{feynman}
\end{tikzpicture}
\ni
\frac{\delta(\textcolor{cut-color}{k_4^2})}{\textcolor{ucp-color}{s\, k_4^2}}
n \left[
  \begin{tikzpicture}
    \ver{e1}{-1,0};
    \ver{e2}{-0.6,0.4};
    \ver{e3}{-0.3,0.4};
    \ver{i1}{-0.6,0};
    \ver{i2}{-0.3,0};
    \ver{i3}{0,0};
    \ver{i4}{0.8,0};
    \ver{co}{$(i3)!0.5!(i4)$};
    \ver{e4}{1.2,0};
    \draw[ultra thick] (e1)--(i1)--(i2)--(i3) to [out=80,in=100,min distance=0.4cm](i4)--(e4);
    \draw[ultra thick,ucp-color](i1)--(i2)--(i3);
    \draw[ultra thick,cut-color](i4)--(e4);
    \draw[ultra thick] (i3) to [out=-80,in=-100,min distance=0.4cm](i4);
    \draw[ultra thick] (e2) -- (i1);
    \draw[ultra thick] (e3)--(i2);
    \draw[ultra thick,dashed,cut-color] (co)--++(90:0.5);
    \draw[ultra thick,dashed,cut-color] (co)--++(-90:0.5);
  \end{tikzpicture}
  \right]
\end{equation}
in which one of the \textcolor{ucp-color}{uncut propagators} is set
equal to an \textcolor{cut-color}{external on-shell propagator}.
Discussions about how to resolve these types of singularities can be
found in Refs.~\cite{Edison:2022jln, Edison:2022smn, Edison:2020uzf,
  Geyer:2015bja, Bern:2012uf} \citepls{Minahan}.  In addition to the
bubble-on-external-leg cuts, there is a related class of cuts in which
an uncut propagator is set to zero by an \emph{internal cut
  condition}, for instance,
\begin{equation}
  \label{eq:other-path}
  \begin{tikzpicture}
    \pgfmathsetmacro{\ri}{1};
    \pgfmathsetmacro{\re}{0.6};
    \pgfmathsetmacro{\ang}{35};
    \begin{feynman}
      \blob{i1}{\ang:\ri};
      \blob{i2}{180-\ang:\ri};
      \blob{i3}{180+\ang:\ri};
      \blob{i4}{-\ang:\ri};
      \blob{i5}{ $(i1)!0.5!(i2)$ };
      \ver{e1}{45:\ri+\re};
      \ver{e2}{135:\ri+\re};
      \ver{e3}{225:\ri+\re};
      \ver{e4}{-45:\ri+\re};

@@

Ignoring tadpoles and BEL graphs, there are 14 cubic 4pt 2-loop graphs (see later picture) which are related to each other via 21 Jacobi relations.
As mentioned in (review section), color-kinematics duality ensures that the numerator of every graph can be expressed in terms of a basis of the double box and penta-triangle,
\begin{equation}
\label{eq:jacobi-basis}
\NLSMBasisDoubleBox \text{ and } \NLSMBasisPentaTriangle .
\end{equation}
For a scalar theory like NLSM the numerator only depends on dot products of momenta.
The basis of Mandelstam invariants for both of these graphs can be taken to be
\begin{equation}
\label{eq:scalar-basis}
\{ k_{13}, k_{23}, k_{15}, k_{16}, k_{25}, k_{26}, k_{35}, k_{36}, k_{55}, k_{56}, k_{66} \},
\end{equation}
where $k_{ij}$ means $k_i \cdot k_j$, $k_5=\ell_1$, and $k_6 = \ell_2$.
%%%%
% Basis for both diagrams
% {d[p[1], p[3]], d[p[1], p[6]], d[p[1], p[7]], d[p[2], p[3]],  d[p[2], p[6]], d[p[2], p[7]], d[p[3], p[6]], d[p[3], p[7]],  d[p[6], p[6]], d[p[6], p[7]], d[p[7], p[7]]}
%%%%
Every vertex in NLSM scales as $k^2$ so the numerator of each diagram scales as $k^{12}$.
At this mass dimension there are 8,008 independent dot products of momenta.
With two basis graphs there is a total of 16,016 ansatz parameters.\footnote{\texttt{Mathematica} is not suited to equations with sixteen thousand parameters so a different method is desirable.
One possibility is to use a solver designed for integral reduction.
However, these solvers typically assume that the matrix has a very low density which is not the case for equations coming from NLSM which typically have a density of a few percent.
While the equations in YM naturally separate into different ``helicity'' sectors (or $e_i \cdot e_j$ and $e_i \cdot k_j$ sectors in $D$ dimensions), there is no such separation of sectors in a scalar theory where any product of momenta can interfere with any other.
The NLSM constraint equations were solved using a custom solver, \texttt{FiniteFieldSolve}, designed to solve large linear systems of arbitrary density.
\texttt{FiniteFieldSolve} will be released publicly shortly.}
Imposing all of the Jacobi relations eliminates all but 4,473 of the original free parameters.
Further imposing graph symmetries reduces the number of free parameters to 1,243.

@@

We follow the same general procedure as outlined above, again
identifying the double-box and penta-triangle as the Jacobi basis
graphs and building a generic ansatz for each of their numerators.
Because we are now considering pure Yang--Mills, each monomial in the
local ansatz must consist of five Lorentz scalar dot products instead
of the six for NLSM, and each monomial must be linear in each of the
four external gluon polarizations.  Thus, the ansatz will be built
from terms of the form
\begin{align}
  \pol_1 \cdot k_2, \quad \pol_1 \cdot k_3, \quad& \pol_1 \cdot \ell_1, \quad \pol_1 \cdot \ell_2, \notag \\
  \pol_2 \cdot k_1, \quad \pol_2 \cdot k_3, \quad& \pol_2 \cdot \ell_1, \quad \pol_2 \cdot \ell_2, \notag \\
  \pol_3 \cdot k_1, \quad \pol_3 \cdot k_2, \quad& \pol_3 \cdot \ell_1, \quad \pol_3 \cdot \ell_2,  \\
  \pol_4 \cdot k_1, \quad \pol_4 \cdot k_2, \quad& \pol_4 \cdot \ell_1, \quad \pol_4 \cdot \ell_2, \notag \\
  \pol_1 \cdot \pol_2, \quad \pol_1 \cdot \pol_3, \quad& \pol_1 \cdot \pol_4, \quad \pol_2 \cdot \pol_3, \notag \\
  \pol_2 \cdot \pol_4, \quad& \pol_3 \cdot \pol_4  \notag
\end{align}
in addition to the purely momentum-dependent terms show in
\cref{eq:scalar-basis}.  In our labeling of the two loop Jacobi-basis
diagrams, shown in \cref{eq:jacobi-basis}, we find that the power
counting bound proposed by Bern, Davies, and Nohle leads to an initial
ansatz with 10,000 terms in the double-box, and 9849 terms for the
penta-triangle.  Without power counting restrictions, both the
double-box and the penta-triangle ansatz have an ansatz with 10,010
terms each.  Imposing diagram automorphisms and maximal-cut gauge
invariance on the two basis diagrams, we reduce the number of terms to
XX for the double-box, and YY for the penta-triangle.  The
power-counting restriction of BDN would further reduce these counts to
2232 and 4072.  In Ref.~\cite{Bern:2015ooa}, the unitarity matching
constraints were imposed using a spanning set of minimal physical
cuts, shown in \cref{fig:ym-spanning}.  Instead of also directly going
to the spanning cuts, we work within the framework of the Method of
Maximal Cuts \citepls{} in order to identify the simplest cut that is
in tension with the kinematic Jacobi relations.
